Analysis of the inference of ratings and rankings on Higher Order Networks with complex topologies

### 2025-09-06 draft/draft.tex
progress

Expression~\ref{eq:1} can be reformulated in the language of discrete calculuas as
\begin{equation}
\label{eq:4}
|f-d_0(w)|
\end{equation}
if the scalar components $f_{ij}$ of the 1-cochain $f=f_{ij}e^{ij}$ are identified with the pairwise comparison socres and the 0-cochain scalar components $w_i$ are identified with the ratings.
In fact, the vector of ratings minimizing expression~\ref{eq:4} is the solution of Eq.~\ref{eq:3} for the case $k=1$.
For this, the set of 0-simplices $K_0$ is identified with a set $V$ of teams on a league, the set of 1-simplices $K_1$ is identified with the set $E$ of links or pairwise comparison of teams in the league.
Additionally, to compute the 2-cochain $u$ in order to obtain the solenoid component of the Hodge decomposition of $f$ from Eq.~\ref{eq:2}, the set of 2-simplices is identified with the set of connected triples in the set of comparisons; i.e.
$$
K_2
=
\{
\{i,j,k\}
:
\{i,j\}\in E,
\{j,k\}\in E,
\{i,k\}\in E
\}.
$$

@@

With each metric, the coefficients
$$
f_{ij}
=
\frac{1}{2}(m_{ij}+m_{ji})
$$
are defined for each $i$ and $j$ in $\{0,1,...,n\}$ such that $i<j$ and where $N=n+1$ is the number of teams in the corresponding league.
Moreover, in order to test the inference method, two types of ground-truth coefficients are also defined.
Namely, the rank-based coefficients $f_{ij}=i-j$ and the tournament points based coefficients $f_{ij}=\pi_j-\pi_i$.
These coefficients correspond to a fully connected network or graph with $N$ nodes since, as already mentioned, all pairs of teams play two matches against each other in each league.
Next, each fully connected graphs is promoted to a simplicial complex composed of 0-simplices labeled by $i$, 1-simplices labeled by $ij$ with $i<j$ and 2-simplices labeled by $ijk$ with $i<j<k$.
Finally, the coefficients $f_{ij}$ are promoted to be the components of a 1-cochain  $f$, where the anti-symmetric relation $f_{ji}=-f_{ij}$ must hold by definition.

>>>>

Expression~\ref{eq:1} can be reformulated in the language of discrete calculuas as
\begin{equation}
\label{eq:4}
|f-d_0(w)|
\end{equation}
if the scalar components $f_{ij}$ of the 1-cochain $f=f_{ij}e^{ij}$ are identified with the pairwise comparison socres and the 0-cochain scalar components $w_i$ are identified with the ratings.
In fact, the vector of ratings minimizing expression~\ref{eq:4} is the solution of Eq.~\ref{eq:3} for the case $k=1$.
For this, the set of 0-simplices $K_0$ is identified with a set $V$ of teams on a league, the set of 1-simplices $K_1$ is identified with the set $E$ of links or pairwise comparison of teams in the league.
Additionally, to compute the 2-cochain $u$ in order to obtain the solenoid component of the Hodge decomposition of $f$ from Eq.~\ref{eq:2}, the set of 2-simplices is identified with the set of connected triples in the set of comparisons; i.e.
$$
K_2
=
\{
\{i,j,k\}
:
\{i,j\}\in E,
\{j,k\}\in E,
\{i,k\}\in E
\}.
$$
In this setting, $K_k=\emptyset$ for  $k>2$.

@@

With each metric, the coefficients
$$
f_{ij}
=
\frac{1}{2}(m_{ij}+m_{ji})
=
- f_{ji}
$$
are defined for each $i$ and $j$ in $\{0,1,...,n\}$ such that $i<j$ and where $N=n+1$ is the number of teams in the corresponding league.
Moreover, in order to test the inference method, two types of ground-truth coefficients are also defined.
Namely, the rank-based coefficients $f_{ij}=i-j$ and the tournament points based coefficients $f_{ij}=\pi_j-\pi_i$.
These coefficients correspond to a fully connected network or graph with $N$ nodes since, as already mentioned, all pairs of teams play two matches against each other in each league.
Next, each fully connected graphs is promoted to a simplicial complex composed of 0-simplices labeled by $i$, 1-simplices labeled by $ij$ with $i<j$ and 2-simplices labeled by $ijk$ with $i<j<k$.
Finally, the coefficients $f_{ij}$ are promoted to be the components of a 1-cochain  $f$, where the anti-symmetric relation $f_{ji}=-f_{ij}$ must hold by definition.
